---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# analysistools

<!-- badges: start -->
[![check-standard](https://github.com/impact-initiatives/analysistools/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/impact-initiatives/analysistools/actions/workflows/check-standard.yaml)
[![Codecov test coverage](https://codecov.io/gh/impact-initiatives/analysistools/branch/main/graph/badge.svg)](https://app.codecov.io/gh/impact-initiatives/analysistools?branch=main)
<!-- badges: end -->

The goal of analysistools is to ...

## Installation

You can install the development version of analysistools from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("impact-initiatives/analysistools")
```

## Example
```{r load}
library(analysistools)
```

This is a basic example which shows you how to calculate the mean:
```{r mean}
somedata <- data.frame(aa = 1:10,
                      bb = rep(c("a","b"),5),
                      weights = rep(c(.5,1.5),5),
                      stratas = rep(c("strata_a", "strata_b"),5))
me_design <- srvyr::as_survey(somedata)
create_analysis_mean(me_design, analysis_var = "aa")
create_analysis_mean(me_design, group_var = "bb", analysis_var = "aa")
me_design_w <- srvyr::as_survey(somedata, weights = weights)
create_analysis_mean(me_design_w, analysis_var = "aa")
create_analysis_mean(me_design_w, group_var = "bb", analysis_var = "aa")

```
This is a basic example which shows you how to calculate the median:

```{r median}
somedata <- data.frame(aa = 1:10,
                      bb = rep(c("a","b"),5),
                      weights = rep(c(.5,1.5),5),
                      stratas = rep(c("strata_a", "strata_b"),5))
me_design <- srvyr::as_survey(somedata)
create_analysis_median(me_design, analysis_var = "aa")
create_analysis_median(me_design, group_var = "bb", analysis_var = "aa")
me_design_w <- srvyr::as_survey(somedata, weights = weights)
create_analysis_median(me_design_w, analysis_var = "aa")
create_analysis_median(me_design_w, group_var = "bb", analysis_var = "aa")
```

This is a basic example which shows you how to calculate the proportion for select one:

```{r proportion}
somedata <- data.frame(groups = sample(c("group_a", "group_b"), size = 100,
                       replace = TRUE),
                       value = sample(c("a", "b", "c"), size = 100, replace = TRUE,
                       prob = c(.6,.4,.1)))

create_analysis_prop_select_one(srvyr::as_survey(somedata, strata = groups),
                                group_var = NA,
                                analysis_var = "value",
                                level = .95)
create_analysis_prop_select_one(srvyr::as_survey(somedata, strata = groups),
                                group_var = "groups",
                                analysis_var = "value",
                                level = .95)
```


